---
title: "Dplyr and GGplot walkthrough"
author: "Sam Edelstein"
date: "February 11, 2016"
output: pdf_document
---

This is a walkthrough of creating a quick report where the requirement includes importing data, cleaning it, and then visualizing it. Some of the processes will be done using base packages in R, but others will use popular packages called dplyr and ggplot2. First, we need to indicate which packages we will be using.

```{r, echo=FALSE}
require(RCurl) #helps read data
library(dplyr) #helps clean data
library(ggplot2) #helps visualize data
```

Import and read data from Google Spreadsheet

```{r}
myCsv <- getURL("https://docs.google.com/spreadsheets/d/1PUZlBLgRGI-6hnDF_hLcn8hmxksIBIp6VZEncgt52wk/pub?output=csv")
RoadRatings <- read.csv(textConnection(myCsv), stringsAsFactors = FALSE)
```

Ratings in the City should only be from 1-10, anything else is a type or inaccurate, so filter out

```{r}
RoadRatings <- filter(RoadRatings, Overall <=10 & Overall >= 1)
```

Count how many road blocks there are per road rating category, then create a data frame

```{r}
RoadsPerRating <- data.frame(table(RoadRatings$Overall))
names(RoadsPerRating) <- c("Rating", "NumRoads") #rename columns so they are easier to understand
RoadsPerRating$Rating <- as.numeric(RoadsPerRating$Rating) #convert rating to numeric data type
```

This step is in preparation for the data visualization. Categorize the road ratings into good, fair, and poor

```{r}
RoadsPerRating$Status <- "Fair"
RoadsPerRating$Status[RoadsPerRating$Rating > 7] <- "Good"
RoadsPerRating$Status[RoadsPerRating$Rating < 6] <- "Poor"
RoadsPerRating$Status <- factor(RoadsPerRating$Status, levels = c("Poor", "Fair", "Good"))
```

Visualize the data using ggplot

```{r}
ggplot(RoadsPerRating, aes(Rating, NumRoads, fill = Status)) + #calls out data and aesthetics that should be used (in aes(x-axis, y-axis, extra information))
  geom_bar(stat="identity") + #indicates we will create a bar chart. "identity" makes the heights of the bars equal the value you are using on the y-axis
  scale_fill_manual(values=c("red", "yellow", "green")) + #determine color of each bar (this is optional but makes sense for good, fair, poor)
  labs(title="Distribution of Roads by Rating", #name the title and axes
       x = "Rating",
       y = "Number of Road Blocks") +
  scale_x_continuous(breaks = c(0:10)) + # make a tick mark at every Rating value
  theme_bw() #take out the background color
```
